---
title: <center><font size="7"><b>Degradation statistics</b></font></center>
subtitle: <center><font size="4"><b>baRulho paper</b> <br>Universidad de Costa Rica</font></center>
author: <center><font size="3"><a href="https://marceloarayasalas.weebly.com/">Marcelo Araya-Salas</a></font></center>
date: <center>"`r Sys.Date()`"</center>
output:
  html_document:
    code_folding: hide
    # css: extra.css
    df_print: tibble
    highlight: pygments  
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

<div class="alert alert-info">

# Purposes

- Estimate degradation metrics on re-recorded signals from playback experiment at Bosque de Tlalpan, Mexico City, 2019

- Explore graphically variation in degradation parameters related to abiotic factors

</div>

&nbsp; 


```{r load packages, echo = FALSE, message = FALSE, warning=FALSE}

# github packages must include user name ("user/package")
# knitr is require for creating html/pdf/word reports
# kableExtra is used to print pretty formatted tables 
# formatR is used for soft-wrapping code
# klippy is used for adding a copy button to each code block
pkgs <- c("remotes", "kableExtra", "knitr", "formatR", "rprojroot", "maRce10/Rraven", "maRce10/warbleR", "maRce10/baRulho", "ggplot2", "tidyr", "viridis", "corrplot", "brms", "ggdist", "cowplot")

# install/ load packages
out <- lapply(pkgs, function(y) {
  
  # get pakage name
  pkg <- strsplit(y, "/")[[1]]
  pkg <- pkg[length(pkg)]
  
  # check if installed, if not then install 
  if (!pkg %in% installed.packages()[,"Package"])  {

      if (grepl("/", y))  remotes::install_github(y, force = TRUE) else
    install.packages(y) 
    }

  # load package
  a <- try(require(pkg, character.only = T), silent = T)

  if (!a) remove.packages(pkg)
  })


```


```{r setup chunks, include = FALSE}

# options to customize chunk outputs
knitr::opts_chunk$set(
  tidy.opts = list(width.cutoff = 65), 
  tidy = TRUE,
  message = FALSE,
  warning = FALSE
 )

```


```{r functions, eval = TRUE, echo = FALSE}

paral <- 22

cols <- viridis(10)[c(3, 8)]

cols2 <- viridis(10)[c(3, 6)]

theme_set(theme_classic(base_size = 17))

# degrad_params <- c(bl.rt = "blur.ratio", sp.bl.rt = "spectrum.blur.ratio", env.cr = "envelope.correlation",  EA = "excess.attenuation.", SPL = "SPL", SNR.t1 = "signal.to.noise.ratio",SNR.t2 = "signal.to.noise.ratio.t2", SPCC = "cross.correlation", TSR = "tail.to.signal.ratio", TNR = "tail.to.noise.ratio", sp.cr = "spectrum.correlation", PC1 = "PC1")

degrad_params <- c(bl.rt = "blur.ratio", sp.bl.rt = "spectrum.blur.ratio", env.cr = "envelope.correlation", EA ="excess.attenuation", SPL = "SPL", SNR.t1 = "signal.to.noise.ratio", SNR.t2 = "signal.to.noise.ratio", SPCC = "cross.correlation",  TSR = "tail.to.signal.ratio", TNR = "tail.to.noise.ratio", sp.cr = "spectrum.correlation", PC1 = "PC1")

cap_fun <- function(x) {
    s <- strsplit(x, " ")[[1]]
    paste(toupper(substring(s, 1, 1)), substring(s, 2),
          sep = "", collapse = " ")
}

```

```{r read data est simulations, eval = FALSE, echo = FALSE}

est.alg.sim.all <- readRDS("./data/raw/extended_sel_tab_tests_simulated_tests_sounds.RDS")

names(est.alg.sim.all)[13] <- "distance" 

est.alg.sim.all$sound.id <- as.character(est.alg.sim.all$template)
```

```{r measure degradation, eval = FALSE, echo = FALSE}

est.alg.sim.all <- set_reference_sounds(est.alg.sim.all)

options(mc.cores = 24, hop.size = 11.6, ovlp = 10, pb = TRUE)

# blur ratio
est.alg.sim.all <- blur_ratio(X = est.alg.sim.all)

# envelope correlation
est.alg.sim.all <- envelope_correlation(est.alg.sim.all)

# excess attenuation
est.alg.sim.all <- excess_attenuation(X = est.alg.sim.all, ovlp = 50, wl = 1024, cores = 24)

# Signal-to-noise ratio
est.alg.sim.all <- signal_to_noise_ratio(est.alg.sim.all, mar = 0.025)

# Cross-correlation
est.alg.sim.all <- spcc(est.alg.sim.all)

# Spectral blur-ratio
est.alg.sim.all <- spectrum_blur_ratio(X = est.alg.sim.all)

# Spectral correlation
est.alg.sim.all <- spectrum_correlation(est.alg.sim.all)

# Tail-to-signal ratio
est.alg.sim.all <- tail_to_signal_ratio(est.alg.sim.all, pb = TRUE,  mar = 0.01)

# habitat attenuation
est.alg.sim.all$environm <- NULL

write.csv(as.data.frame(est.alg.sim.all), "./data/processed/barulho_degradation_metrics.csv", row.names = FALSE)
# 
# degrad_df$SNR <- sapply(degrad_df$sound.files, function(x) est.alg.sim.all$signal.to.noise.ratio[est.alg.sim.all$sound.files == x])

np <- noise_profile(X = est.alg.sim.all, mar = 0.01, noise.ref = "adjacent")

np$amp <- np$amp * -1 

write.csv(np, "./data/processed/noise_profiles.csv", row.names = FALSE)

```


```{r snr, eval = FALSE}

degrad_df <- read.csv("./data/processed/barulho_degradation_metrics.csv", stringsAsFactors = FALSE)

degrad_df$signal.to.noise.ratio <- NA
# Signal-to-noise ratio
degrad_df$signal.to.noise.ratio[est.alg.sim.all$end - est.alg.sim.all$start > 0.15] <- signal_to_noise_ratio(est.alg.sim.all[est.alg.sim.all$end - est.alg.sim.all$start > 0.15, ], mar = 0.025, pb = TRUE, cores = paral, type = 1)$signal.to.noise.ratio

degrad_df$signal.to.noise.ratio[est.alg.sim.all$end - est.alg.sim.all$start < 0.15] <- signal_to_noise_ratio(est.alg.sim.all[est.alg.sim.all$end - est.alg.sim.all$start < 0.15, ], mar = 0.025 / 2, pb = TRUE, cores = paral, type = 1)$signal.to.noise.ratio

```

```{r read data degradation, eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, include=FALSE}

degrad_df <- read.csv("./data/processed/barulho_degradation_metrics.csv", stringsAsFactors = FALSE)

master_metadata <- read.csv("./data/raw/consolidated.master.sf_selection_table.csv")

# keep only simulated data (exluding data from a student)
master_metadata <- master_metadata[2:961, ]

# infer original frequency
possb_freqs <- seq(0.5, 10, length.out = 20)

master_metadata$freq <- sapply(1:nrow(master_metadata), function(x)
  possb_freqs[which.min(abs((master_metadata$top.freq[x] + master_metadata$bottom.freq[x]) / 2 - possb_freqs))]  
)

freq_contour <- freq_ts(master_metadata, length.out = 20, cores =22, img = FALSE, path = "./data/raw/master_sound_file", pb = FALSE)

freq_contour$mean.freq <- rowMeans(freq_contour[, 3:ncol(freq_contour)])

freq_contour$orig_sound_file <- master_metadata$orig.sound.file

degrad_df$frequency <- sapply(1:nrow(degrad_df), function(x)
  freq_contour$mean.freq[freq_contour$orig_sound_file == degrad_df$sound.id[x]]
  )

degrad_df$sim.frequency <- sapply(1:nrow(degrad_df), function(x)
  master_metadata$freq[master_metadata$orig.sound.file == degrad_df$sound.id[x]]
  )


degrad_df$habitat_structure <- ifelse(degrad_df$Vegetacion == "abierta", "open", "closed")

degrad_df <- separate(degrad_df, col = sound.id, into = c("Duration", "Freq_modulation", "Amp_modulation", "Harmonicity", "delete"), remove = FALSE, sep = "_")

degrad_df$Duration <- ifelse(degrad_df$Duration == 0.1, "short", "long")
degrad_df$`Freq_modulation` <- ifelse(degrad_df$`Freq_modulation` == "BB", "fm", "no_fm")
degrad_df$Harmonicity <- gsub(".wav", "", degrad_df$Harmonicity)
degrad_df$delete <- NULL

degrad_df$categ <- paste(degrad_df$habitat_structure, as.numeric(degrad_df$sim.freq), sep = "-")

```

```{r add PCA, out.width = "100%"}

pca_degrad_params <- c("blur.ratio", "spectrum.blur.ratio", "envelope.correlation", "excess.attenuation", "signal.to.noise.ratio", "cross.correlation", "tail.to.signal.ratio", "tail.to.noise.ratio", "spectrum.correlation")

comp.cases <- complete.cases(degrad_df[,names(degrad_df) %in% pca_degrad_params])

pca <- prcomp(degrad_df[comp.cases, names(degrad_df) %in% pca_degrad_params], scale. = TRUE)

# add to data
degrad_df$PC1 <- NA
degrad_df$PC1[comp.cases] <- pca$x[, 1]

# plot rotation values by PC
pca_rot <- as.data.frame(pca$rotation[, 1:4])

pca_rot_stck <- stack(pca_rot)

pca_rot_stck$variable <- rownames(pca_rot)
pca_rot_stck$Sign <- ifelse(pca_rot_stck$values > 0, "Positive", "Negative")
pca_rot_stck$rotation <- abs(pca_rot_stck$values)

ggplot(pca_rot_stck, aes(x = variable, y = rotation, fill = Sign)) +
    geom_col() + 
    coord_flip() +
    scale_fill_viridis_d(alpha = 0.7, begin = 0.3, end = 0.8) +
    facet_wrap(~ ind)
```


```{r correlation among metrics, out.width = "120%"}

cormat <- cor(degrad_df[, degrad_params], use = "pairwise.complete.obs")

rownames(cormat) <- colnames(cormat) <- names(degrad_params)

cols_corr <- colorRampPalette(c("white", "white", viridis(4, direction = -1)))(10)

cp <- corrplot.mixed(cormat, tl.cex = 0.7,
               upper.col = cols_corr,
               lower.col = cols_corr,
               order = "hclust",
               lower = "number", 
               upper = "ellipse",
               tl.col = "black")

# sort parameters as in clusters for cross correlation
degrad_params <- degrad_params[match(rownames(cp$corr), names(degrad_params))]

```

# Change with distance by frequency
```{r plots distance, eval = TRUE, echo = FALSE, fig.width= 10, results='asis'}

plts <- lapply(degrad_params, function(x){
  
  # remove distance 1 m for those that compared against 1m
  X <- if (x %in%  c("blur.ratio", "spectrum.blur.ratio", "PC1", "excess.attenuation.fixed", "envelope.correlation", "cross.correlation", "spectrum.correlation", "excess.attenuation.sw"))
  degrad_df[degrad_df$distance != 1, ] else degrad_df

  
  form <- as.formula(paste(x, "~ sim.frequency + habitat_structure + distance"))
  
  agg <- aggregate(form, data = X, mean)
  
  names(agg)[ncol(agg)] <- "variable"
  
  agg$sd <- aggregate(form, data = X, sd)[, x] 
  
  plt1 <- ggplot(agg, aes(x = distance, y = variable, col = habitat_structure, group = habitat_structure)) +
    geom_line(position = position_dodge(0.25)) +
    geom_point(position = position_dodge(0.25))+
    scale_color_viridis_d(alpha = 0.7, begin = 0.2, end = 0.8) +
    geom_errorbar(aes(ymin = variable - sd, 
                      ymax = variable + sd), width = .2,
                   position = position_dodge(0.25)) +
    facet_wrap(~sim.frequency) + 
    labs(x = "Distance (m)", y = gsub("\\.", " ", x)) +
    scale_x_continuous(breaks=sort(unique(X$distance))) +
    theme_classic() 
  
  X$by.m <- X[, x] * 10 / X$distance
  
   agg <- aggregate(by.m ~ sim.frequency + habitat_structure + distance, data = X, mean)
  
  names(agg)[ncol(agg)] <- "variable"
  
  agg$sd <- aggregate(by.m ~ sim.frequency + habitat_structure + distance, data = X, sd)[, "by.m"] 
  
  plt2 <- ggplot(agg, aes(x = distance, y = variable, col = habitat_structure, group = habitat_structure)) +
    geom_line(position = position_dodge(0.25)) +
    geom_point(position = position_dodge(0.25))+
    scale_color_viridis_d(alpha = 0.7, begin = 0.2, end = 0.8) +
    geom_errorbar(aes(ymin = variable - sd, 
                      ymax = variable + sd), width = .2,
                   position = position_dodge(0.25)) +
    facet_wrap(~sim.frequency) + 
    labs(x = "Distance (m)", y = paste0(gsub("\\.", " ", x), " / 10 m")) +
    scale_x_continuous(breaks=sort(unique(X$distance))) +
    theme_classic() 
  return(list(plt1, plt2))
    
})

names(plts) <- degrad_params

for(i in 1:length(plts)){
  
  cat(paste('<font size="2"> ', gsub("\\.", " ", cap_fun(names(plts)[i])), "</font>"))

  cat("<br>")
  
  print(plts[[i]][[1]])
  print(plts[[i]][[2]])
  
  cat("&nbsp;")
  
  }
  
```


# Habitat structure
```{r plots hab, eval = TRUE, echo = FALSE, fig.width= 10}

plts <- lapply(degrad_params, function(x){
  
  # remove distance 1 m for those that compared against 1m
  # X <- if (x %in%  c("blur.ratio", "spectrum.blur.ratio", "PC1", "excess.attenuation.ma", "envelope.correlation", "cross.correlation", "spectrum.correlation", "excess.attenuation.da"))
  # degrad_df[degrad_df$distance != 1, ] else degrad_df
  
  X <- degrad_df[degrad_df$distance != 1, ]
  
  form <- as.formula(paste(x, "~ sim.frequency + habitat_structure + distance"))
  
  agg <- aggregate(form, data = X, mean)
  
  names(agg)[ncol(agg)] <- "variable"
  
  agg$sd <- aggregate(form, data = X, sd)[, x] 
  
  plt <- ggplot(agg, aes(x = sim.frequency, y = variable, col = habitat_structure, group = habitat_structure)) +
    geom_line(position = position_dodge(0.25)) +
    geom_point(position = position_dodge(0.25))+
    scale_color_viridis_d(alpha = 0.7, begin = 0.2, end = 0.8) +
    geom_errorbar(aes(ymin = variable - sd, 
                      ymax = variable + sd), width = .2,
                   position = position_dodge(0.25)) +
    facet_wrap(~distance) + 
    labs(x = "Frequency (kHz)", y = gsub("\\.", " ", x)) +
    theme_classic() 
  
  return(plt)
    
})

names(plts) <- degrad_params

for(i in 1:length(plts)){
  
  print(names(plts)[i])

  print(plts[[i]])
  }
  
```

# Frequency modulation and habitat structure
```{r plots fm, eval = TRUE, echo = FALSE, fig.height = 10, fig.width= 10}


plts <- lapply(degrad_params, function(x){
  
  # remove distance 1 m for those that compared against 1m
  # X <- if (x %in%  c("blur.ratio", "spectrum.blur.ratio", "PC1", "excess.attenuation.fixed", "envelope.correlation", "cross.correlation", "spectrum.correlation", "excess.attenuation.sw"))
  # degrad_df[degrad_df$distance != 1, ] else degrad_df
  X <- degrad_df[degrad_df$distance != 1, ]
  
  form <- as.formula(paste(x, "~ sim.frequency + habitat_structure + distance + Freq_modulation"))
  
  agg <- aggregate(form, data = X, mean)
  
  names(agg)[ncol(agg)] <- "variable"
  
  agg$sd <- aggregate(form, data = X, sd)[, x] 
  
  plt <- ggplot(agg, aes(x = sim.frequency, y = variable, col = Freq_modulation, group =  Freq_modulation)) +
    geom_line(position = position_dodge(0.25)) +
    geom_point(position = position_dodge(0.25))+
    scale_color_viridis_d(alpha = 0.7, begin = 0.2, end = 0.8) +
    geom_errorbar(aes(ymin = variable - sd, 
                      ymax = variable + sd), width = .2,
                   position = position_dodge(0.25)) +
    facet_wrap(~distance + habitat_structure, ncol = 2) + 
    labs(x = "Frequency (kHz)", y = gsub("\\.", " ", x)) +
    theme_classic() 
  
  return(plt)
    
})

names(plts) <- degrad_params

for(i in 1:length(plts)){
  
  print(names(plts)[i])

  print(plts[[i]])
  }
  
  
```


# Amplitude modulation and habitat structure
```{r plots amp, eval = TRUE, echo = FALSE, fig.height = 10, fig.width= 10}

plts <- lapply(degrad_params, function(x){
  
   # remove distance 1 m for those that compared against 1m
  # X <- if (x %in%  c("blur.ratio", "spectrum.blur.ratio", "PC1", "excess.attenuation.fixed", "envelope.correlation", "cross.correlation", "spectrum.correlation", "excess.attenuation.sw"))
  # degrad_df[degrad_df$distance != 1, ] else degrad_df
  X <- degrad_df[degrad_df$distance != 1, ]
  
  form <- as.formula(paste(x, "~ sim.frequency + habitat_structure + distance + Amp_modulation"))
  
  agg <- aggregate(form, data = X, mean)
  
  names(agg)[ncol(agg)] <- "variable"
  
  agg$sd <- aggregate(form, data = X, sd)[, x] 
  
  plt <- ggplot(agg, aes(x = sim.frequency, y = variable, col = Amp_modulation, group =  Amp_modulation)) +
    geom_line(position = position_dodge(0.25)) +
    geom_point(position = position_dodge(0.25))+
     scale_color_viridis_d(alpha = 0.7, begin = 0.2, end = 0.8) +
    geom_errorbar(aes(ymin = variable - sd, 
                      ymax = variable + sd), width = .2,
                   position = position_dodge(0.25)) +
    facet_wrap(~distance + habitat_structure, ncol = 2) + 
    labs(x = "Frequency (kHz)", y = gsub("\\.", " ", x)) +
    theme_classic() 
  
  return(plt)
    
})

names(plts) <- degrad_params

for(i in 1:length(plts)){
  
  print(names(plts)[i])

  print(plts[[i]])
  }
  
```

# Harmonicity and habitat structure
```{r plots harm, eval = TRUE, echo = FALSE, fig.height = 10, fig.width= 10}


plts <- lapply(degrad_params, function(x){
  
  # remove distance 1 m for those that compared against 1m
  # X <- if (x %in%  c("blur.ratio", "spectrum.blur.ratio", "PC1", "excess.attenuation.fixed", "envelope.correlation", "cross.correlation", "spectrum.correlation", "excess.attenuation.sw"))
  # degrad_df[degrad_df$distance != 1, ] else degrad_df
  X <- degrad_df[degrad_df$distance != 1, ]
  
  form <- as.formula(paste(x, "~ sim.frequency + habitat_structure + distance + Harmonicity"))
  
  agg <- aggregate(form, data = X, mean)
  
  names(agg)[ncol(agg)] <- "variable"
  
  agg$sd <- aggregate(form, data = X, sd)[, x] 
  
  plt <- ggplot(agg, aes(x = sim.frequency, y = variable, col = Harmonicity, group =  Harmonicity)) +
    geom_line(position = position_dodge(0.25)) +
    geom_point(position = position_dodge(0.25))+
     scale_color_viridis_d(alpha = 0.7, begin = 0.2, end = 0.8) +
    geom_errorbar(aes(ymin = variable - sd, 
                      ymax = variable + sd), width = .2,
                   position = position_dodge(0.25)) +
    facet_wrap(~distance + habitat_structure, ncol = 2) + 
    labs(x = "Frequency (kHz)", y = gsub("\\.", " ", x)) +
    theme_classic() 
  
  return(plt)
    
})

names(plts) <- degrad_params

for(i in 1:length(plts)){
  
  print(names(plts)[i])

  print(plts[[i]])
  }
  
```

# Duration and habitat structure
```{r plots dur, eval = TRUE, echo = FALSE, fig.height = 10, fig.width= 10}

plts <- lapply(degrad_params, function(x){
  
  # remove distance 1 m for those that compared against 1m
  # X <- if (x %in%  c("blur.ratio", "spectrum.blur.ratio", "PC1", "excess.attenuation.fixed", "envelope.correlation", "cross.correlation", "spectrum.correlation", "excess.attenuation.sw"))
  # degrad_df[degrad_df$distance != 1, ] else degrad_df
    X <- degrad_df[degrad_df$distance != 1, ]

  form <- as.formula(paste(x, "~ sim.frequency + habitat_structure + distance + Duration"))
  
  agg <- aggregate(form, data = X, mean)
  
  names(agg)[ncol(agg)] <- "variable"
  
  agg$sd <- aggregate(form, data = X, sd)[, x] 
  
  plt <- ggplot(agg, aes(x = sim.frequency, y = variable, col = Duration, group =  Duration)) +
    geom_line(position = position_dodge(0.25)) +
    geom_point(position = position_dodge(0.25))+
     scale_color_viridis_d(alpha = 0.7, begin = 0.2, end = 0.8) +
    geom_errorbar(aes(ymin = variable - sd, 
                      ymax = variable + sd), width = .2,
                   position = position_dodge(0.25)) +
    facet_wrap(~distance + habitat_structure, ncol = 2) + 
    labs(x = "Frequency (kHz)", y = gsub("\\.", " ", x)) +
    theme_classic() 
  
  return(plt)
    
})

names(plts) <- degrad_params

for(i in 1:length(plts)){
  
  print(names(plts)[i])

  print(plts[[i]])
  }
  
```

# Temperature
```{r plots temp, eval = TRUE, echo = FALSE, fig.height = 3, fig.width = 10}

qnts <- quantile(degrad_df$Temperatura, c(0.10, 0.90))

sub.degrad_df <- degrad_df[degrad_df$Temperatura < qnts[1] | degrad_df$Temperatura > qnts[2], ]

# sub.degrad_df <- sub.degrad_df[sub.degrad_df$sim.frequency <= 2 | sub.degrad_df$sim.frequency >= 8.5, ]

sub.degrad_df$temp.cat <- ifelse(sub.degrad_df$Temperatura < qnts[1], "< 24 deg", "> 28 deg")  

# sub.degrad_df$sim.frequency.cat <- ifelse(sub.degrad_df$sim.frequency <= 2, "< 2", ">8")  



plts <- lapply(degrad_params, function(x){
  
  if (x == "spectral.blur.ratio") sub.degrad_df <- sub.degrad_df[!is.na(sub.degrad_df$spectral.blur.ratio), ]

  form <- as.formula(paste(x, "~ sim.frequency + habitat_structure + temp.cat"))
  
  agg <- aggregate(form, data = sub.degrad_df, mean)
  
  names(agg)[ncol(agg)] <- "variable"
  
  agg$sd <- aggregate(form, data = sub.degrad_df, sd)[, x] 
  
  plt <- ggplot(agg, aes(x = sim.frequency, y = variable, col = interaction(temp.cat, habitat_structure), group =  interaction(temp.cat, habitat_structure))) +
    geom_line(position = position_dodge(0.25)) +
    geom_point(position = position_dodge(0.25))+
     scale_color_viridis_d(alpha = 0.7, begin = 0.2, end = 0.8) +
    geom_errorbar(aes(ymin = variable - sd, 
                      ymax = variable + sd), width = .2,
                   position = position_dodge(0.25)) +
    # facet_wrap(~distance, ncol = 2) + 
    labs(x = "Frequency (kHz)", y = gsub("\\.", " ", x)) +
    theme_classic() 
  
  return(plt)
    
})

names(plts) <- degrad_params

for(i in 1:length(plts)){
  
  print(names(plts)[i])

  print(plts[[i]])
  }
  
```


<div class="alert alert-success">

&nbsp; 

# Takeaways


</div>

&nbsp;

<!-- '---' adds a gray vertical line -->

---

&nbsp; 
 
 <!-- add packages used, system details and versions  -->
 
<font size="4">Session information</font>

```{r session info, echo=F}

sessionInfo()

```


